Leadership across cultures and contexts: an exploratory analysis

### 2020-05-11 leadership_across_cultures_contexts.Rmd
A few minor edits

Within benefit/cost dimensions, there was relatively more evidence for mating and territory benefits than costs for both leaders and followers, and equal levels of evidence for protection from harm vs. risk of harm. For leaders compared to followers, there were noticably higher odds ratios for inclusive fitness benefits, material resources, and social services. See Figure \@ref(fig:benefitcostratio).

@@

We then fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting the `r nrow(feature_models_aic)` models whose AIC values were at least two points less than the intercept-only models (only Cultural conformity was dropped). As with our leader dimension models, we then dropped each factor variable in turn and compared the change in AIC. Large increases indicated that including the factor substantially improved fit. See Figure \@ref(fig:heatmapdrop1)A. We then compared the levels of each factor that improved fit for each feature variable. See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).

@@

The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership. See Figure \@ref(fig:heatmapgroups). These results underscore recent trends in evoltuionary approaches to leadership which emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].

@@

@@

We therefore coded all benefits and costs of both leadership and followership to better understand which benefits and costs are most commonly associated with leading and following. Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high risk-return strategy (see Figure \@ref(fig:varsupportcb)).

The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources. Followers were not reported to receive mating benefits, relative to leaders or other types benefits. Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy. For both leaders and followers, individuals stood to lose what they were able to gain. The major difference between leaders and followers lay in leaders' opportunities for increased social status and mating opportunities, relative to followers. See Figure \@ref(fig:varsupportcb).

>>>>

Within benefit/cost dimensions, there was relatively more evidence for mating and territory benefits than costs for both leaders and followers, and equal levels of evidence for protection from harm vs. risk of harm. For leaders compared to followers, there were noticably higher odds ratios for mating, inclusive fitness, material resources, and social services. See Figure \@ref(fig:benefitcostratio).

@@

We then fit separate logistic mixed effects models of each of the `r nrow(feature_models)` features as a function of group context, subsistence strategy, continental region, and leader sex, with random effects for author nested within culture, comparing these to intercept-only models, and selecting the `r nrow(feature_models_aic)` models whose AIC values were at least two points less than the intercept-only models (only Cultural conformity was dropped). As with our leader dimension models, we then dropped each factor variable in turn and compared the change in AIC. The large increases for group context and leader sex indicated that including these factors substantially improved fit. See Figure \@ref(fig:heatmapdrop1)B. We then compared the levels of each factor that improved fit for each feature variable. Female leaders were associated with social, reproductive, and material success, and male leaders with prosociality. Miltary groups were associated the Organize feature, religious groups with Prosociality and Competencies, and and kin groups with Strategizing and Prosocial competencies. See Figures S\@ref(fig:heatmapfeaturesex), and S\@ref(fig:heatmapfeaturegroups).

@@

The clearest pattern to emerge from analyses of systematic variation is that, not too surprisingly, many dimensions of leadership are context-specific: aggressiveness and bravery were associated with military leadership, supernatural qualities and ritual functions were associated with religious leadership, organizing cooperation and being experience or accomplished were associated with economic-group leadership, and providing counsel and older age were associated with kin-group leadership. See Figure \@ref(fig:heatmapgroups). These results underscore recent trends in evoltuionary approaches to leadership that emphasize the context-dependency of leader emergence and evolutionary selection pressures [@smith_leadership_2020; @smith_leadership_2016; @hagen_leadership_2019].

@@

## The rarity of some important leadership dimensions

@@

Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported, suggesting leadership to be a high risk, high return strategy (see Figure \@ref(fig:varsupportcb)). The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources. Followers were not reported to receive mating benefits, relative to leaders or other types benefits. Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy. For both leaders and followers, individuals stood to lose what they were able to gain. The major difference between leaders and followers lay in leaders' opportunities for increased social status and mating opportunities, relative to followers. See Figure \@ref(fig:varsupportcb).
